Digital economy in the UK: a multi-scalar story of the diffusion of web technologies

### 2023-03-21 paper/submission1/tranos2023-IT074305.qmd
paper first draft

---
title: Digital economy in the UK: a multi-scalar story of the diffusion of web technologies
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliation: University of Bristol and The Alan Turing Institute
    correspondingauthor: true
    footnote: 1
address:
  - code: University of Bristol and The Alan Turing Institute
    address: University of Bristol, School of Geographical Sciences, Bristol, UK; The Alan Turing Institute, London, UK
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  This paper maps the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. The data and geolocation strategy allow to capture these processes at a very granular spatial scale. The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies. Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.

keywords: 
  - keyword1
  - keyword2
journal: "An awesome journal"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---
# Introduction {#sec:1}

@@

<!-- key concepts -->
@bento2015spatial The concept of absorptive capacity has also been
applied in the TIS context in terms of the con-struction of capacity
needed to transfer and implement a new technology in the receiving
country (cf.Van Alphen, 2011), with repercussions for the nature of
system functions

@@

<!-- @das2022diffusion -->
<!-- Epidemic approach: The approach emphasizes that the adoption of a -->
<!-- technology spreads like an infection among the potential adopters. By -->
<!-- taking a homogeneous group of potential adopters as an assumption, it -->
<!-- explains the adoption of any innovation. -->

@@

<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->
<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->
<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->

@@

<!-- the three geographic diffusion processes are (1) purely contagious where -->
<!-- distance or adjacency is the absolute controlling factor, (2) purely -->
<!-- hierarchical where size or urban position in the central place hierarchy -->
<!-- is the absolute controlling factor, or (3) where the location of change -->
<!-- is purely random. These are the pure or ideal forms of the diffusion -->
<!-- process; all real diffusion processes are the result of a combination of -->
<!-- these extremes. -->

@@

<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->
<!-- are considered as the three pillars of the theory of technological -->
<!-- change and Schumpeter is considered as the father of the study of -->
<!-- diffusion -->

@@

<!-- @leibowicz2016representing -->
<!-- An illustrative empirically based theory for technology -->
<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->
<!-- (Grubler, 1990). With respect to a specific technology, it divides -->
<!-- regions into three groups: core, rim, and periphery. The core is where -->
<!-- the technology is invented, subject to experimentation, and first -->
<!-- deployed commercially. To develop the technology, the core must have -->
<!-- both sufficient innovative ability and economic motivation to make the -->
<!-- necessary resource commitment. As the technology spreads from the core -->
<!-- to the rim and eventually to the periphery, later adopters have the -->
<!-- advantage of benefiting from the experiences of earlier adopters via -->
<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->
<!-- point, the technology has likely been refined and costs have likely -->
<!-- declined significantly. Schmidt's Law claims that, as a result, -->
<!-- diffusion proceeds at a faster rate in regions which adopt the -->
<!-- technology later. However, diffusion tends to be less pervasive in -->
<!-- regions where adoption is delayed, ultimately saturating at a lower -->
<!-- extent than in regions where adoption occurred earlier. This is often -->
<!-- attributed to a lack of proper institutions and infrastructures in later -->
<!-- adopters, markets whichmay not be as favorable to a technology, or the -->
<!-- availability of yet newer options that invite technological -->
<!-- leapfrogging. -->

@@

<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->
<!-- diffusion is the process by which an innovation is communicated through -->
<!-- certain communication channels over time among the members of a social -->
<!-- system. An innovation may be an idea, object or practice which is -->
<!-- perceived as new by an individual or another unit of adoption (e.g., an -->
<!-- organization) . Communication via communication channels refers to how -->
<!-- knowledge about an innovation is transferred between the members of a -->
<!-- social system. Communication takes time, and is therefore an important -->
<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->
<!-- diffusion studies agree that an innovation does not instantly diffuse -->
<!-- into a social system. Instead a typical time pattern of diffusion along -->
<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->
<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->
<!-- confirmed by a large number of studies· of primitive preindustrial -->
<!-- societies. -->

@@

<!-- grubler1990rise -->
<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->
<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->
<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->
<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->
<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->
<!-- and diffusion eventually ceases. -->

<!-- grubler1990rise -->
<!-- Thus, three main characteristics define the spatial diffusion process: -->
<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->
<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->
<!-- diffusion proceeds from initial centers to secondary ones. Within the -->
<!-- context of urban systems (central places) this implies a hierarchical -->
<!-- spread of an innovation from larger to smaller centers, and from the -->
<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->
<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->
<!-- (Brown, 1981) nearby rather than far-away locations. -->

@@

<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->
<!-- starts from innovative centers ('core´) and disseminates, through a -->
<!-- hierarchy of subcenters, to the periphery. This process is sequential -->
<!-- rather than simultaneous, and tends to accelerate from core to -->
<!-- periphery, where diffusion reaches a lower intensity than in the core -->
<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->
<!-- spatial acceleration has been identified in marketing research and is -->
<!-- known as the "lead-lag effect" (Peres et al., 2010). -->

@@

<!-- methods -->
@wilson201281 Logistic growth describes an initial period of gradual
diffusion as a technology is introduced as a new commercial application,
moving then through a rapid, exponential growth phase, before slowing
and eventually saturating [@grubler1999dynamics]. The substitution of
incumbent technologies by new competitors leads to subsequent decline
and eventual obsolescence.

@@

<!-- previous studies -->
@beardsell1999spatial To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992

>>>>

---
title: "Digital economy in the UK: a multi-scalar story of the diffusion of web technologies"
author:
  - name: Emmanouil Tranos
    email: e.tranos@bristol.ac.uk
    affiliations: 
      - id: some-tech
        name: University of Bristol and The Alan Turing Institute
        department: 
        address: 
        city: 
        state: State
        postal-code: UK
    corresponding: e.tranos@bristol.ac.uk
address:
  - attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
  This paper maps the participation in the digital economy and its evolution in the UK over space and time. Most of the existing economic geography literature which dealt with the spatiality of the internet employed supply-side measures, such as infrastructural capacity, in order to understand the geography of the digital economy and its potential spatial economic effects. Useful as these approaches might have been, they cannot capture the micro-processes and the characteristics of the individual online behaviour. Using large volumes of archived and geolocated web content, this paper models the diffusion of web technologies over space and time in the UK. The data and geolocation strategy allow to capture these processes at a very granular spatial scale. The modelling approach, which is based on simple spatial analytical methods and on the estimation of diffusion curves at various scales, enables to depict the role of geography and other cognitive factors which drove the diffusion of web technologies. Although the focus is on a recent historical period -- 1996-2012 -- the results of the analysis depict diffusion mechanisms which can be very useful in understanding the evolutionary patterns of the adoption of other newer technologies.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
editor: 
  markdown: 
    wrap: 72
---

# Introduction {#sec:1}

@@

<!-- key concepts -->

@bento2015spatial The concept of absorptive capacity has also been
applied in the TIS context in terms of the con-struction of capacity
needed to transfer and implement a new technology in the receiving
country (cf.Van Alphen, 2011), with repercussions for the nature of
system functions

@@

<!-- @das2022diffusion -->

<!-- Epidemic approach: The approach emphasizes that the adoption of a -->

<!-- technology spreads like an infection among the potential adopters. By -->

<!-- taking a homogeneous group of potential adopters as an assumption, it -->

<!-- explains the adoption of any innovation. -->

@@

<!-- Rogers3, that divides adopters as: (I.) Innovators: first 2.5%, (II.) -->

<!-- Early adopters: next 13.5%, (III.) Early Majority: following 34%, (IV.) -->

<!-- Late Majority: next 34%, and the rest as (V.) Laggards -->

@@

<!-- the three geographic diffusion processes are (1) purely contagious where -->

<!-- distance or adjacency is the absolute controlling factor, (2) purely -->

<!-- hierarchical where size or urban position in the central place hierarchy -->

<!-- is the absolute controlling factor, or (3) where the location of change -->

<!-- is purely random. These are the pure or ideal forms of the diffusion -->

<!-- process; all real diffusion processes are the result of a combination of -->

<!-- these extremes. -->

@@

<!-- @das2022diffusion [not good paper] Invention, innovation and diffusion -->

<!-- are considered as the three pillars of the theory of technological -->

<!-- change and Schumpeter is considered as the father of the study of -->

<!-- diffusion -->

@@

<!-- @leibowicz2016representing -->

<!-- An illustrative empirically based theory for technology -->

<!-- diffusion across multiple regions is suggested by "Schmidt's Law" -->

<!-- (Grubler, 1990). With respect to a specific technology, it divides -->

<!-- regions into three groups: core, rim, and periphery. The core is where -->

<!-- the technology is invented, subject to experimentation, and first -->

<!-- deployed commercially. To develop the technology, the core must have -->

<!-- both sufficient innovative ability and economic motivation to make the -->

<!-- necessary resource commitment. As the technology spreads from the core -->

<!-- to the rim and eventually to the periphery, later adopters have the -->

<!-- advantage of benefiting from the experiences of earlier adopters via -->

<!-- knowledge spillover effects (Verdolini and Galeotti, 2011). At this -->

<!-- point, the technology has likely been refined and costs have likely -->

<!-- declined significantly. Schmidt's Law claims that, as a result, -->

<!-- diffusion proceeds at a faster rate in regions which adopt the -->

<!-- technology later. However, diffusion tends to be less pervasive in -->

<!-- regions where adoption is delayed, ultimately saturating at a lower -->

<!-- extent than in regions where adoption occurred earlier. This is often -->

<!-- attributed to a lack of proper institutions and infrastructures in later -->

<!-- adopters, markets whichmay not be as favorable to a technology, or the -->

<!-- availability of yet newer options that invite technological -->

<!-- leapfrogging. -->

@@

<!-- @grubler1990rise Following the definition given by Rogers, 1983, -->

<!-- diffusion is the process by which an innovation is communicated through -->

<!-- certain communication channels over time among the members of a social -->

<!-- system. An innovation may be an idea, object or practice which is -->

<!-- perceived as new by an individual or another unit of adoption (e.g., an -->

<!-- organization) . Communication via communication channels refers to how -->

<!-- knowledge about an innovation is transferred between the members of a -->

<!-- social system. Communication takes time, and is therefore an important -->

<!-- aspect of innovation diffusion. In fact, all theoretical and empirical -->

<!-- diffusion studies agree that an innovation does not instantly diffuse -->

<!-- into a social system. Instead a typical time pattern of diffusion along -->

<!-- a S-shaped trajectory seems to be the rule. The S-shaped pattern of -->

<!-- diffusion appears to be a basic anthropologic phenomenon, as it is also -->

<!-- confirmed by a large number of studies· of primitive preindustrial -->

<!-- societies. -->

@@

<!-- grubler1990rise -->

<!-- Spatial innovation diffusion in Hagerstrand's model consists of three -->

<!-- stages. Stage 1: Local concentrations of initial acceptances (initial -->

<!-- agglomeration) . Stage 2: Radial dissemination outward from the initial -->

<!-- agglomeration, accompanied by the rise of secondary agglomerations. -->

<!-- Stage 3: The condensing and saturation stage, in which filling-in occurs -->

<!-- and diffusion eventually ceases. -->

<!-- grubler1990rise -->

<!-- Thus, three main characteristics define the spatial diffusion process: -->

<!-- (1) As in temporal diffusion models, an S-shaped pattern in the -->

<!-- cumulative level of adoption. (2) A "hierarchy effect" by which -->

<!-- diffusion proceeds from initial centers to secondary ones. Within the -->

<!-- context of urban systems (central places) this implies a hierarchical -->

<!-- spread of an innovation from larger to smaller centers, and from the -->

<!-- centers to the surrounding hinterland. (3) A "neighborhood effect" Le., -->

<!-- diffusion proceeds outwards from innovation centers, first "hitting" -->

<!-- (Brown, 1981) nearby rather than far-away locations. -->

@@

<!-- Hagerstrand (1968) was the first to demonstrate that diffusion typically -->

<!-- starts from innovative centers ('core´) and disseminates, through a -->

<!-- hierarchy of subcenters, to the periphery. This process is sequential -->

<!-- rather than simultaneous, and tends to accelerate from core to -->

<!-- periphery, where diffusion reaches a lower intensity than in the core -->

<!-- (Hagerstrand, 1968; Grubler, 1990; Morrill, 2005). A similar effect of -->

<!-- spatial acceleration has been identified in marketing research and is -->

<!-- known as the "lead-lag effect" (Peres et al., 2010). -->

@@

<!-- methods -->

@wilson201281 Logistic growth describes an initial period of gradual
diffusion as a technology is introduced as a new commercial application,
moving then through a rapid, exponential growth phase, before slowing
and eventually saturating [@grubler1999dynamics]. The substitution of
incumbent technologies by new competitors leads to subsequent decline
and eventual obsolescence.

@@

<!-- previous studies -->

@beardsell1999spatial To characterize evolution of the computer
industry, we examine the distribution of relative employment across
cities in 1977 and how that distribution changes over time to 1992
